---
title: "Project 3: STAT302package Tutorial"
author: "Alisha Luo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{STAT302package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ggplot2)
library(reshape)
```

```{r setup}
library(STAT302package)
```


# Introduction
This package contains all functions developed in the course STAT 302. Each 
function demonstrates an inference/prediction method and works with the 
gapminder data. 
## Installation and Use
To download the STAT302package package, use the code below
```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("alishaluo/STAT302package")
library(STAT302package)
devtools::install_github("alishaluo/STAT302package", build_vignette = TRUE, build_opts = c())
library(STAT302package)
# Use this to view the vignette in the corncob HTML help
help(package = "STAT302package", help_type = "html")
# Use this to view the vignette as an isolated HTML file
utils::browseVignettes(package = "STAT302package")
```

# my_t.test function
```{r}
two_side_test <- my_t_test(my_gapminder$lifeExp, alternative = "two.sided", mu = 60)
two_side_test
less_test <- my_t_test(my_gapminder$lifeExp, alternative = "less", mu = 60)
greater_test <- my_t_test(my_gapminder$lifeExp, alternative = "greater", mu = 60)
as.data.frame(cbind(two_side_test$p_value, less_test$p_value,
                    greater_test$p_value))
```

If the life expectancy is an average of 60 years, we cannot conclude that 
9.32% of samples will not equal 60, thus failing to reject the null hypothesis. 
If the life expectancy is an average of 60 years, only 4.66% of samples will
be less than 60. 
If the life expectancy is an average of 60 years, only 4.66% of samples will
be greater than 60. 

# my_lm function
```{r}
test <- my_lm(formula = lifeExp ~ gdpPercap + continent, data = my_gapminder)
test
my_coef <- test[, 1]
my_matrix <- model.matrix(lifeExp ~ gdpPercap + continent, data = my_gapminder)
y_hat <- my_matrix %*% as.matrix(my_coef)

my_data <- data.frame("Actual" = my_gapminder$lifeExp, "Fitted" = y_hat, 
                      "Continent" = my_gapminder$continent)
ggplot(my_data, aes(x = Actual, y = Fitted, color = Continent)) + 
  geom_point() + 
  theme_bw(base_size = 15) +
  geom_abline(slope = 1, intercept = 0, col = "black") +
  labs(x = "Fitted values", y = "Actual values", title = "Actual vs. Fitted") +
  theme(plot.title = element_text(hjust = 0.5))
```

The null hypothesis would be that there is no effect gdpPercap has on life
expectancy. Since the p-value is so low, assuming that gdp per capita does not
have an effect on life expectancy, we can say that 8.55e-71 percent of samples 
will observe gdp per capita having an effect on life expectancy, thus rejecting
the null hypothesis. 

By looking at the graph, we can see that Europe and Oceania were predicted 
pretty well, while for a continent like Africa, we predicted the life 
expectancy to be somewhere from 35 to 75, but the actual life expectancy ranged
closer to 50. 

# my_knn_cv function
```{r}
set.seed(123)
my_output_err <- rep(NA, 10)
my_train_err <- rep(NA, 10)
  
for(i in 1:10) {
  my_output <- my_knn_cv(train = my_gapminder[, c(4, 6)], 
                         cl = my_gapminder$continent, k_nn = i, k_cv = 5)
  train_err <- mean(my_output$class != my_gapminder$continent)
  my_output_err[i] <- (my_output$CV_error)
  my_train_err[i] <- train_err
}
as.data.frame(cbind(my_output_err, my_train_err))
```

```{r, warning= FALSE}
set.seed(321)
cv_mse_2 <- rep(NA, 30)
cv_mse_5 <- rep(NA, 30)
cv_mse_10 <- rep(NA, 30)
for(i in 1:30) {
  cv_mse_2[i] <- my_rf_cv(k = 2)
  cv_mse_5[i] <- my_rf_cv(k = 5)
  cv_mse_10[i] <- my_rf_cv(k = 10)
}
cv_mse <- as.data.frame(cbind(cv_mse_2, cv_mse_5, cv_mse_10))
cv_mse_melt <- melt(cv_mse)
ggplot(cv_mse_melt, aes(factor(variable), value)) +
  geom_boxplot() +
  labs(x = "Number of Folds", y = "CV Error") +
  scale_x_discrete(labels = c("2", "5", "10"))
avg <- rep(NA, 3)
std_dev <- rep(NA, 3)
for(i in 1:length(cv_mse)) {
  avg[i] <- mean(cv_mse[, i])
  std_dev[i] <- sd(cv_mse[, i])
}
as.data.frame(cbind(avg, std_dev))
```

In both the graph and table, we can see that as the number of folds increases, 
the cross validation error increase, we also see the standard deviation
decrease. 
